The non-random assembly of network motifs in  \nplant-pollinator networks

### 2023-01-08 Manuscript/Main_text/Main_text_Journal_of_Animal_Ecol.Rmd
Graphical abstract and word doc for word count
edits

Despite the obvious usefulness of functional groups in summarising species ecological roles (e.g., similar reproductive biology for plants and behaviour for pollinators), delimiting species into functional groups is often not trivial. For instance, taxonomic diversity and trait availability limit our ability to properly represent the wide range of reproductive strategies [@dellinger2020] and different clustering analysis can lead to distinct outcomes (e.g., k-means and hierarchical clustering). We divided floral visitors on the taxonomic rank level because it adequately represented pollinator ecological roles without the tremendous effort of compiling trait data for over 1,000 species of floral visitors. Indeed, further groupings with traits could help reveal greater relevance of trait-matching processes on indirect interactions. Moreover, we analysed network subgraphs focusing exclusively on presence or absence of interactions, assuming that all interactions are equally important. However, further work should investigate how to account for quantitative measurements of plant-pollinator interactions. Consistent with previous work [@simmons2019; @simmons2020], we characterised network motifs by using null models only constrained for network connectance. Although more complex null models are difficult to interpret [@Ulrich2013], they can potentially allow us to separate the mechanisms by which motifs emerge. However, we found that null models with more constraints generated binary structures that were mostly identical to the original ones. In addition, we observed that connectance, nestedness and degree distribution were significantly correlated (data not shown). All that highlights the difficulties to disentangle the potential sources that generate over-representation of network subgraphs and the need to investigate them. At the same time, although network subgraphs can capture both direct and indirect interactions, the reduction of network complexity into these different building blocks inevitably convey the loss of information [@simmons2019], for example indirect interactions from large cascading effects. Finally, despite the wide geographical coverage of our plant-pollinator dataset, there are major gaps for some regions of the world with a clear bias towards Europe, America and islands which have long interested ecologists [@macarthur1963; @warren2015]. These biases are also found in global datasets of networks, where the data is highly fragmented across space [@poisot2021] and more efforts are needed to capture global patterns of plant-pollinator interactions.

@@

>>>>

Despite the obvious usefulness of functional groups in summarising species ecological roles (e.g., similar reproductive biology for plants and behaviour for pollinators), delimiting species into functional groups is often not trivial. For instance, taxonomic diversity and trait availability limit our ability to properly represent the wide range of reproductive strategies [@dellinger2020] and different clustering analysis can lead to distinct outcomes (e.g., k-means and hierarchical clustering). We divided floral visitors on the taxonomic rank level because it adequately represented pollinator ecological roles without the tremendous effort of compiling trait data for over 1,000 species of floral visitors. Indeed, further groupings with traits could help reveal greater relevance of trait-matching processes on indirect interactions. Moreover, we analysed network subgraphs focusing exclusively on presence or absence of interactions, assuming that all interactions are equally important. However, further work should investigate how to account for quantitative measurements of plant-pollinator interactions. Consistent with previous work [@simmons2019; @simmons2020], we characterised network motifs by using null models only constrained for network connectance. Although more complex null models are difficult to interpret [@Ulrich2013], they can potentially allow us to separate the mechanisms by which motifs emerge. However, we found that null models with more constraints generated binary structures that were mostly identical to the original ones. In addition, we observed that connectance, nestedness and degree distribution were significantly correlated (data not shown). All that highlights the difficulties to disentangle the potential sources that generate over-representation of network subgraphs and the need to investigate them. At the same time, although network subgraphs can capture both direct and indirect interactions, the reduction of network complexity into these different building blocks inevitably convey the loss of information [@simmons2019]. For example, network subgraphs neglect indirect effects emerging from large cascading effects that propagate across them. Finally, despite the wide geographical coverage of our plant-pollinator dataset, there are major gaps for some regions of the world with a clear bias towards Europe, America and islands which have long interested ecologists [@macarthur1963; @warren2015]. These biases are also found in global datasets of networks, where the data is highly fragmented across space [@poisot2021] and more efforts are needed to capture global patterns of plant-pollinator interactions.

@@

# Conflict of interest declaration

The authors declare no conflict of interest.
